A comparison of simple models for urban morphogenesis

### 2020-08-30 Docs/Papers/MesoBenchmark/paper/paper.tex
.

@@

Following the so-called ``first law of geography'', entities in space have interaction patterns which can be described with spatial interaction models, including the gravity model. \cite{li2019singularity}

@@

@@

\cite{10.1371/journal.pone.0203516}
% modify: max density?

@@

\cite{makse1998modeling}

The method to generate a spatial field exhibiting long range correlations was introduced for problems in physics by \cite{makse1996method}.

@@

Finally, to provide some kind of null model to understand the advantages of each approach compared to a simple description of population distribution, we also include urban forms generated as kernel mixtures. We consider in particular exponential mixtures \cite{anas1998urban}, written as

@@

Quantitative measures of urban form

@@

% TODO examples for each model

@@

@@

@@

\section*{Acknowledgments}

>>>>

We study and compare four different models of urban morphogenesis. We consider a population grid of size $N = W\times H$ (not necessarily square), each cell being characterized by its population $P_i$ with $1\leq i \leq N$.

@@

@@

Following the so-called ``first law of geography'', entities in space have interaction patterns which can be described with spatial interaction models \cite{fotheringham1989spatial}, including the gravity model. \cite{li2019singularity} proposed an urban growth model including this process within an iterative growth with population aggregation processes, extending a more simple model introduced by \cite{rybski2013distance}. We generalize this model by adding (i) a hierarchy parameter regarding population aggregation and (ii) seeding multiple initial sites to allow the emergence of polycentric urban forms.

Formally, an initial grid is seeded with $P_0^{(G)}$ sites with population 1, randomly selected. Then, iteratively, one unit of population is added to each cell at each time step with a probability proportional to 
\begin{equation}
p_i \propto \frac{\sum_{j\neq i} P_j^{\gamma_P^{(G)}} \cdot d_{ij}^{- \gamma_D^{(G)}}}{\sum d_{ij}^{- \gamma_D^{(G)}}}
\end{equation}

where probabilities are rescaled such that the cell with the larger value has a probability of $g^{(G)}$. This last parameter allows modifying the speed of growth. The model is stopped when a total population $P_M^{(G)}$ is reached.

@@

% modify: max density? - NO

In his attempt to understand embryogenesis, Alan Turing proposed to use chemical partial differential equations (PDEs) to model morphogenesis, introducing the nowadays famous reaction-diffusion equations \cite{turing1990chemical}. In such systems, chemical substances react together and diffuse in space, leading to the emergence of complex geometrical patterns. The concept of morphogenesis has been since well used in urban studies \cite{raimbault2018co}, but very few models have actually implemented reaction-diffusion equations, \cite{bonin2014modelisation} being a notable exception. \cite{10.1371/journal.pone.0203516} proposes to capture the fundamental processes of agglomeration economies (positive externalities) leading to aggregation and of congestion (negative externalities) leading to sprawl, as an ``aggregation-diffusion'' model of urban morphogenesis. The model yields indeed in certain limits reaction-diffusion PDEs. Formally, starting from an empty grid, $N^{(R)}$ units of population are added at each time step, and attributed independently to cells with a probability proportional to $P_i^{\alpha^{(R)}}$ (probabilities are rescaled to obtain a probability distribution over all cells). Population is then diffused in space $d^{(R)}$ times with a strength $\beta^{(R)}$. The model is stopped when a maximum population $P_M^{(R)}$ is reached.

@@

The first two models presented are iterative and can in theory be used dynamically. Other approaches, closer to procedural modeling \cite{parish2001procedural}, do not simulate the progressive growth of population. They can however capture processes at play in the growth of urban form. The correlated percolation model described by \cite{makse1998modeling} integrates for example clustering processes in cities. A method to generate a spatial field exhibiting long range correlations was introduced for problems in physics by \cite{makse1996method}. It is combined to a monocentric density profile in \cite{makse1998modeling} to produce urban forms. In practice, a correlated field $p_i$ is generated by (i) generating a random spatial field; (ii) compute its spatial Fourier transform; (iii) introduce a correlation by multiplying it with a spectral density function with a power-law exponent $\alpha^{(C)}$; (iv) retrieve a long-range correlated spatial field by taking the inverse Fourier transform. This field is combined to a density field $\rho_i$ to determine a binary value for the cell: it is populated if $p_i > \theta_i$ with $\rho_i = \int_{-\infty}^{\theta_i} d\mathbb{P}(p_i)$. We generalize the initial model by taking a polycentric density field with

\begin{equation}
\rho_i (\vec{x}) = \sum_{j=1}^{n^{(C)}} \exp\left(\lVert\vec{x} - \vec{x}_j\rVert / r_j\right)
\label{eq:kernelmixture}
\end{equation}

where the kernel centers are chosen at random and kernel sizes $r_j$ are taken such that kernel populations follow a rank-size law of hierarchy $\beta^{(C)}$ and the largest kernel has a fixed size $r_0^{(C)}$.

@@

Finally, to provide some kind of null model to understand the advantages of each approach compared to a simple description of population distribution, we also include urban forms generated as kernel mixtures. We consider in particular exponential mixtures \cite{anas1998urban}, where population density is written as previously in Eq.~\ref{eq:kernelmixture}. Parameters for this model are the number of kernels $n^{(E)}$, the rank-size hierarchy $\alpha^{(E)}$, the size of the largest kernel $r_0^{E}$. Contrary to previous models in which total population had an influence by controlling the speed of growth, density can here be rescaled arbitrarily (morphological indicators used are not changed through rescaling, see below), and we set the maximal density for one kernel to one.

@@

Quantitative measures of urban form are multiple and depend on the scale considered \cite{zhang2005metrics}. \cite{raimbault2019generating} for example introduces measures for buildings at the district scale. The field of Landscape Ecology has its own metrics similar to urban form measures \cite{10.1371/journal.pone.0225734}. For the scale we consider and considering population distribution only, metrics of urban form have been proposed for example to quantify sprawl \cite{tsai2005quantifying}. These can be related to fractal approaches to urban form \cite{chen2011derivation}. The effective dimension when applied to real cities is reasonably low \cite{schwarz2010urban}, and a few complementary indicators can be used. We thus follow \cite{10.1371/journal.pone.0203516} and consider urban form measures which are: (i) Moran index to capture spatial autocorrelation and the existence of centers; (ii) average distance between individuals which captures a level of aggregation; (iii) distribution entropy (aspatial) to capture the uniformity of the distribution; and (iv) rank-size slope which captures the hierarchy of population distribution.

@@

@@

\subsection*{Generated urban forms}

% examples for each model

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig3.png}
	\caption{{\bf .}\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig4.png}
	\caption{{\bf .}\label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%%

@@

%\section*{Acknowledgments}
